DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-05-17 2015CompleteDerivation/completeDerivation.tex
Minor improvements

On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$. By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$. This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differential, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities. Under arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.

For the Jacobian to exist the transformation between state variables has to be differentiable. Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold. $\mathcal{S}$ has a differentiable structure in the sense that it is on that structure that densities are defined. In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differential changes of state variable can be detected because they are incompatible with state-variable-independent densities.

If $\rho_\mathcal{c}$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This means that we cannot change physical units of all the variables as we please: they are part of a unit system. For example, changing units of position will also change units of momentum. Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units. An element of $\mathcal{S}$ will be a geometrical object in such space. For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined. As the particle state represent configurations of infinitesimal parts, it will be a local object in $\mathcal{Q}$, defined around a point. As we saw before, it cannot be a point itself as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant. It cannot be a vector $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant. A covector $k(q) dq$ is instead fine as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. In other words: a covector is the simplest local geometrical object that allows coordinate independent densities. We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.

@@

This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$. The canonical one-form $\theta=k(q) dq$ represents the geometrical object we associate with each particle state, which is not a point in $\mathcal{Q}$. The canonical two-form $\omega=\Sigma \, dq^i \wedge dk_i$ is the metric that quantifies the possibilities described by two given state variables. Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.

>>>>

On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$. For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$. By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$. This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using. This is inconsistent with our previous definition. Moreover if the transformation is not differential, the density is not even well-defined. This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities. Under an arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.

For the Jacobian to exist the transformation between state variables has to be differentiable. Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold. $\mathcal{S}$ has a differentiable structure in the sense that the distributions are only defined on a set of set state variables that are linked by differential transformations. In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differential changes of state variable can be detected because they are incompatible with state-variable-independent densities.

If $\rho_\mathcal{c}$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please. If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary. This means that we cannot change physical units of all the variables as we please: they are part of a unit system. For example, changing units of position will also change units of momentum. Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units. An element of $\mathcal{S}$ will be a geometrical object in such space. For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined. As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$. As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant. They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant. They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$. In other words: a covector is the simplest local geometrical object that allows coordinate independent densities. We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.

@@

This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$. The canonical one-form $\theta=k dq$ represents the geometrical object we associate with each particle state. The canonical two-form $\omega=\Sigma \, dq^i \wedge dp_i$ is the metric that quantifies the possibilities described by two given state variables. Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.
